Accessing the nucleon transverse structure in inclusive deep inelastic scattering

### 2017-07-28 1607 - JMC for g2/Paper/Reply_to_Referee.txt

​We thank the reviewer for the thoughtful comments and questions about our work, that we have carefully considered and incorporated in the revised version of our paper. In the following, we comment upon each point raised by the reviewer, and refer to the appropriate place in the text where major changes have occurred. (As the numbering of the equations has changed slightly in the revision, in our answers we refer to the new numbering.) We are confident the new version will satisfy the Referee, but will be glad to make further improvements if needed.

@@

 > 1. The new term, the last term in Eq. (21), has the non-perturbative
 > mass scale M_q and the (perturbative) current quark mass m_q come in
 > on the same footing. Have the authors tried to verify their
 > conclusions in an explicit calculation? For instance, perhaps
 > calculating the correlator \Xi from (4) in some non-perturbative model
 > may help illustrate the fact that M_q - m_q is in fact non-zero.

We have indeed checked the behavior of the Mq term in a model with a pseudoscalar Yukawa interaction modeling the quark fragmentation process (see Ref. [28]). This clearly shows that Mq>mq. We added a comment to this effect at the end of Section II together with a revised discussion of the estimate of the magnitude of Mq, see next.

@@

 > 2. If I understood the arguments above Eq. (9) correctly,
 > strictly-speaking the conclusion there is that M_q < O (100 MeV), not
 > equal to O (100 MeV) as Eq. (9) has. In fact, {\bar J}_2 can be very
 > small as far as one can tell, leading to M_q = m_q. Some
 > clarifications would help here.

The Referee is right in noticing that nothing a priori prevents Mq to become smaller than ~100 MeV. Our estimate was based on the fact that one can expect the two-particle continuum to open up at (mq+m_pi)~m_pi~Lambda_QCD, and therefore the value of mu averaged over the J2 distribution should be around our originally quoted value. We have revised this estimate taking also into account that a quark is in fact not a physical asymptotic state, and Mq could be smaller than that, and now quote Mq~10-100 MeV.

Note that having exactly Mq=mq would imply the existence of a so far unknown symmetry or otherwise general constraint that would be extremely interesting on its own. Although we judge this unlikely, our paper may, if nothing else, ​​​encourage people to prove us wrong and discover a new feature of QCD.

The modified discussion, incorporating also comment #1, can be found at the end of Section II.

@@

 > 3. The sum rule (31) raises the question of whether the integral on
 > its right-hand side is finite. It is hard for me to imagine a
 > realistic scenario in which it is. In perturbation theory unpolarized
 > TMDs at small x scale as (1/x)^{1+\epsilon_1}, where \epsilon_1 is an
 > order-\alpha_s positive perturbative correction, while 1 in the power
 > is ultimately determined by the gluon spin =1. Some polarized TMDs,
 > such as helicity, scale as (1/x)^{\epsilon_2} at small x with
 > \epsilon_2 an order-\sqrt{\alpha_s} positive perturbative correction,
 > as found e.g. in Ref. [41]. Note that neither of those behaviors gives
 > a finite right-hand side of (31). The scenario considered by the
 > authors in Eq. (32) corresponds to negative \epsilon_2: while I know
 > of no theorem requiring \epsilon_2 > 0, I also know of no perturbative
 > calculation where \epsilon_2 < 0 either. (Furthermore, in N=4 SYM
 > theory, where the unpolarized intercept is known to a very high order
 > in perturbation theory, the series in \alpha_s changes sign only every
 > other term, see Eq. (5.1) in 1504.06640.) Saturation effects at very
 > low x can at most effectively put \epsilon_2 =0, still not making the
 > right-hand side of (31) finite... So, as a friendly colleague I urge
 > the authors to rethink their conclusions very carefully.

We agree with the referee, and in fact our original discussion was somewhat confusing. We have extended the scope of the discussion and reviewed some perturbative and resummed results on the small x behavior of h1. The small-x physics of the h1 function in comparison to g1 is indeed an interesting topic of research, that can now also be accessed experimentally ​in a new way thanks to our extended WW and BC relations.

The modified discussion can be found in the paragraph above Eq.(32).

@@

 > 4. Here are some minor corrections:
 >
 > -- Ref. [22] has 3 authors, but only two are mentioned explicitly in
 >    the text three lines above Eq. (1).
 >
 > -- The kinematics of the diagrams in Fig. 1 is not defined: e.g., is
 >    the proton moving in the light-cone "+" or "-" direction? This is
 >    needed to better understand the discussion between Eqs. (6) and
 >    (7).
 >
 > -- The equation in the text two lines above Eq. (9) should have m_q^2
 >    (instead of m_q) inside the delta-function.

We thank the referee also for this level of attention to our work, that has indeed helped us improving the article.​ The kinematic has been described in the caption of Figure 1, and the other two typos corrected.​

### 2017-09-11 1607 - JMC for g2/Paper/JMC_dis.tex
Corrected to published version; as resubmitted to arXiv.

The tensor charge is a fundamental property of the nucleon that is at
present poorly constrained but of fundamental importance, not the least because its knowledge can also be used to put constraints on searches for physics beyond the Standard Model~\cite{Cirigliano:2013xha,Bhattacharya:2015esa,Courtoy:2015haa}.
The tensor charge has been estimated in lattice QCD (see, \eg,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn}), but
only limited information is available from direct measurements. Its experimental extraction requires first of all flavor-separated measurements of the so-called transversity parton distribution function,
denoted by $h_1^q(x)$ (see Ref.~\cite{Barone:2001sp} for a review and
Refs.~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa} for the most recent
extractions). Secondly, one needs to perform flavor-by-flavor integrals of these, that correspond to the contribution of a parton flavor $q$ to the tensor charge.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions $A_i$ and $B_i$ through a Lorentz covariant Dirac decomposition that utilizes the vectors $l$ and $n_+$, 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ ,
\label{e:jetexpansion}
\end{equation} 
where $\Lambda$ is an arbitrary scale, introduced for power counting purposes.
In this parametrization, no terms proportional to $\gamma_5$ enter because of parity invariance. Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes only at twist-4 order and will not be considered further in this paper. We focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
    + {\cal O}(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the cut quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one (or, in other words, the interacting rather than the free quark fields).
The $A_1$ and $A_2$ terms can be interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu\,\id + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ can be interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a). The $J_i$ are the spectral functions of the
quark propagator, also called ``jet functions'' in
\cite{Accardi:2008ne}, and can be interpreted as current-jet mass distributions. As a consequence of positivity constraints and CPT invariance, these satisfy \cite{D'Hoker:2004aa,Romao:2013aa,Weinberg:1995mt}
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqs.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.

It is important to notice that, while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass
$\mj < \int d\mu^2 \mu \, J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we expect that $J_2(\mu^2) = Z \delta(\mu^2-m_q^2) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$, with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects, indicating $M_q=O(\Lambda_{QCD})$. However, in a dynamical confinement scenario, the spectral function $J_2$ needs not be positive definite \cite{Roberts:2016vyn} and we therefore estimate
$
  M_q\sim 10-100 \text{\ MeV}.
$
An experimental measurement of $M_q$ is anyway possible, as we discuss in Section~\ref{sec:moments}, and could shed some light on the confinement mechanism. We have also explicitly verified that $M_q>m_q$ in a model where quark fragmentation is simulated by a Yukawa pseudoscalar quark-meson interaction, already utilized, {\it e.g.}, in Ref.~\cite{Bacchetta:2001di}.

@@

Finally, studying and classifying all the contributions of jet correlators to single and double hadron production in electron-positron annihilation events will open up a rich phenomenology. Measurements in the asymptotically large $Q^2$ regime will provide access to the integral of the $J_1$ jet function, i.e., to the jet-mass parameter $\mj$, and therefore (in conjunction with precise measurements or lattice QCD calculations of the first $h_1$ moment) also of the target's tensor charge through the modified ETL sum rule. 
Equally interesting is the possibility to experimentally measure, at finite values of $Q^2$, the momentum dependence of the jet functions $J_1$ and $J_2$, that enter structure functions integrated only up to $\mu^2=Q^2(1/x_B-1)$~\cite{Accardi:2008ne}. In other words, it may become possible to experimentally access also the quark's spectral function itself.

>>>>

The tensor charge is a fundamental property of the nucleon that is at
present poorly constrained but of fundamental importance, not the least because its knowledge can also be used to put constraints on searches for physics beyond the Standard Model~\cite{Cirigliano:2013xha,Bhattacharya:2015esa,Courtoy:2015haa,Tomalak:2017owk,Yamanaka:2017mef}.
The tensor charge has been estimated in lattice QCD (see, \eg,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn}), but
only limited information is available from direct measurements. Its experimental extraction requires first of all flavor-separated measurements of the so-called transversity parton distribution function,
denoted by $h_1^q(x)$ (see Ref.~\cite{Barone:2001sp} for a review and
Refs.~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa} for the most recent
extractions). Secondly, one needs to perform flavor-by-flavor integrals of these, that correspond to the contribution of a parton flavor $q$ to the tensor charge.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions $A_i$ and $B_i$ through a Lorentz covariant Dirac decomposition that utilizes the vectors $l$ and $n_+$, 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ ,
\label{e:jetexpansion}
\end{equation} 
where $\Lambda$ is an arbitrary scale, introduced for power counting purposes.
In this parametrization, no terms proportional to $\gamma_5$ enter because of parity invariance. Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes only at twist-4 order and will not be considered further in this paper. We focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,\id + A_2(l^2)\,\lslash 
    + {\cal O}(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the cut quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one (or, in other words, the interacting rather than the free quark fields).
The $A_1$ and $A_2$ terms can be interpreted in terms of the spectral
representation of the cut quark propagator (see, e.g., Sec.~6.3 of
\cite{D'Hoker:2004aa} and Sec.~2.7.2 of \cite{Romao:2013aa}),
\begin{align} 
  \Xi(l) =  
  \int d \sigma^2 \big[ J_1 (\sigma^2)\,\sigma\,\id + J_2 (\sigma^2)\,\lslash \big] \,
  \delta(l^2 -\sigma^2) \ ,
\label{e:jetspectral}
\end{align}
where $\sigma^2$ can be interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a). The $J_i$ are the spectral functions of the
quark propagator, also called ``jet functions'' in
\cite{Accardi:2008ne}, and can be interpreted as current-jet mass distributions. As a consequence of positivity constraints and CPT invariance, these satisfy \cite{D'Hoker:2004aa,Romao:2013aa,Weinberg:1995mt}
\begin{align}
  J_2(\sigma^2) \geq J_1(\sigma^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\sigma^2 J_2(\sigma^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqs.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\sigma^2 \frac{\sigma}{\Lambda} J_1(\sigma^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\sigma^2 J_2(\sigma^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.

It is important to notice that, while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass
$\mj < \int d\sigma^2 \sigma \, J_2(\sigma^2)$ is dynamically determined. From the analytic properties of the spectral functions we expect that $J_2(\sigma^2) = Z \delta(\sigma^2-m_q^2) + \bar J_2(\sigma^2) \theta (\sigma^2-m_\pi^2)$, with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects, indicating $M_q=O(\Lambda_{QCD})$. However, in a dynamical confinement scenario, the spectral function $J_2$ needs not be positive definite \cite{Roberts:2016vyn} and we therefore estimate
$
  M_q\sim 10-100 \text{\ MeV}.
$
An experimental measurement of $M_q$ is anyway possible, as we discuss in Section~\ref{sec:moments}, and could shed some light on the confinement mechanism. We have also explicitly verified that $M_q>m_q$ in a model where quark fragmentation is simulated by a Yukawa pseudoscalar quark-meson interaction, already utilized, {\it e.g.}, in Ref.~\cite{Bacchetta:2001di}.

@@

Finally, studying and classifying all the contributions of jet correlators to single and double hadron production in electron-positron annihilation events will open up a rich phenomenology. Measurements in the asymptotically large $Q^2$ regime will provide access to the integral of the $J_1$ jet function, i.e., to the jet-mass parameter $\mj$, and therefore (in conjunction with precise measurements or lattice QCD calculations of the first $h_1$ moment) also of the target's tensor charge through the modified ETL sum rule. 
Equally interesting is the possibility to experimentally measure, at finite values of $Q^2$, the momentum dependence of the jet functions $J_1$ and $J_2$, that enter structure functions integrated only up to $\sigma^2=Q^2(1/x_B-1)$~\cite{Accardi:2008ne}. In other words, it may become possible to experimentally access also the quark's spectral function itself.
